\Large \bf Evaluating eBPF as a Platform for Congestion Control Algorithm

### 2024-12-12 paper/sections/abstract.tex
paper: revise

\documentclass[../main.tex]{subfiles}
\begin{document}
This paper seeks to explore the viability of developing Congestion Control Algorithms (CCAs) via eBPF. Developing new CCAs has traditionally been a time-consuming and error-prone process. Creating these algorithms directly within the Linux kernel demands a high level of kernel programming expertise while still leaving room for bugs that can crash the system. Although implementing CCAs as kernel modules has traditionally been an accepted alternative (https://pbg.cs.illinois.edu/papers/jay18pcc-kernel.pdf), it still runs the risk of introducing instability and crashes. Recent research has implemented a Congestion Control Plane (CCP) \cite{ccp} that allows CCA developers to implement CCAs in userspace, eliminating the need to run experimental software directly in the kernel. This approach involves a carefully designed and safety-checked datapath API; however, the Linux kernel datapath API still relies on a kernel module. Simultaneously, developing CCAs in eBPF \cite{ebpf_cca} has emerged as another promising approach, leveraging the eBPF verifier to ensure safety before the code ever runs in the kernel. In this paper, we compare these various approaches -- kernel-based, eBPF-based, and userspace (via CCP) -- along with a new hybrid userspace eBPF-based approach (ebpfccp) in terms of behavioral correctness and performance overhead.
\end{document}

>>>>

\documentclass[../main.tex]{subfiles}
\begin{document}
This paper seeks to explore the viability of developing Congestion Control Algorithms (CCAs) via eBPF. Developing new CCAs has traditionally been a time-consuming and error-prone process. Creating these algorithms directly within the Linux kernel demands a high level of kernel programming expertise while still leaving room for bugs that can crash the system. Although implementing CCAs as kernel modules has traditionally been an accepted alternative \cite{Jay2018APK}, it still runs the risk of introducing instability and crashes. Recent research has implemented a Congestion Control Plane (CCP) \cite{ccp} that allows CCA developers to implement CCAs in user space, eliminating the need to run experimental software directly in the kernel. This approach involves a carefully designed and safety-checked datapath API; however, the Linux kernel datapath API still relies on a kernel module. Simultaneously, developing CCAs in eBPF \cite{ebpf_cca} has emerged as another promising approach, leveraging the eBPF verifier to ensure safety before the code ever runs in the kernel. In this paper, we compare these various approaches -- kernel-based, eBPF-based, and user space (via CCP) -- along with a new hybrid user space eBPF-based approach (ebpfccp) in terms of behavioral correctness and performance overhead.
\end{document}

### 2024-12-12 paper/sections/conclusion.tex
paper: revise

\documentclass[../main.tex]{subfiles}
\begin{document}
Conclusion
\end{document}

>>>>

\documentclass[../main.tex]{subfiles}
\begin{document}
In this paper, we introduced an eBPF-based platform called eBPFCCP, which enables the implementation of user space CCAs. We then compared eBPFCCP with the native Linux kernel TCP stack, a pure eBPF implementation, and CCP. Our evaluations focused on the behavioral correctness and performance impact of these implementations. Through our behavior evaluations, we demonstrated that CCAs implemented using alternative platforms achieve the desired behavior. Additionally, we showed that eBPF is a suitable platform for developing CCAs natively, with negligible performance overhead compared to the kernel’s native implementation. While our implementation of eBPFCCP had a slightly higher performance overhead, it still achieved a substantial percentage of the full kernel implementation’s value. We believe that eBPFCCP can be valuable as an alternative CCP datapath that offers safety guarantees. Overall, the data presented here demonstrates that using eBPF-related CCA platforms not only provides numerous benefits, including enhanced development speed and safety, but also has the potential for long-term deployment beyond research stages. 
\end{document}

### 2024-12-12 paper/sections/evaluation.tex
paper: revise

\subsection{Setup and Method}
We believe that an alternative implementation of an existing CCA can be evaluated using two metrics. First, whether or not the behavior of the implementation matches (to a reasonable extent) existing implementations of the same CCA. Secondly, the performance costs of the implementation when compared to other implementations. We included two sets of benchmarks to measure this: a behavioral benchmark aimed to verify the "correctness" of different implementations (TCP Cubic should behave similarly no matter where it is implemented, as indicated by metrics such as bitrate/congestion window size over time), and a performance benchmark aimed to measure the CPU overhead of different implementation approaches. A CCA implementation with smaller CPU overhead would logically lead to a higher maximum bandwidth in an ideal, infinite-speed connection where the CPU running the CCA is the bottleneck. \\
Our benchmarks are ran with iperf3 \cite{iperf}. For each trial, we loop through every CCA we are testing and run iperf3 for 15 seconds. At the same time, we run a Python script that loops infinitely and adds 1 to an accumulator at every loop. Once the 15 seconds are up, the script then outputs the final number into a file, a higher number suggesting that the script was able to access more CPU cycles. We ran 20 trials to ensure statistical reliability. Our testing script also included various parameters that can be tuned, including trial duration, reporting intervals, and the metric to be used while generating graphs for the behavioral benchmark. \\
After all the trials, we average the behavioral metric across the trials and plot the average for each CCA through time as the behavioral benchmark. We then plot the CPU performances for each CCA across trials on a boxplot. \\

In order to make sure we are correctly benchmarking the performance impact of CCAs rather than miscellneous confounding factors, we ran the benchmark on an out-of-the-box Debian image on a VM. To ensure the CPU benchmark and the CCA are running on the same CPU, our VM is set up to contain only one core. We also decided to run our benchmarks on various implementations of TCP Cubic due to its wide availability across platforms.

@@

The results of behavioral benchmarks are shown in Figures~\ref{fig:cca-behavior-cwnd},~\ref{fig:cca-behavior-bitrate}, and~\ref{fig:cca-behavior-rtt}. From the graph, we can see that the behavior of the same Congestion Control Algorithm remains similar across multiple platforms of integration. These results gives us confidence on the correctness of all the implementation platforms involved, and allows us to proceed to performance benchmarking. The results of CPU benchmarks are shown in Figure \ref{fig:cca-performance}. The eBPF implementation of TCP Cubic achieved impressive CPU performance compared to the built-in Linux kernel networking stack implementation. CCP-Cubic from the CCP Project \cite{generic-cong-avoid} combined with the kernel module + netlink CCP datapath achieved slightly lower performance, where its median is at x percent of the full kernel implementation value. Our implementation of eBPFCCP ranked lower on the performance chart, but was still able to achieve y percent of the full kernel implementation value. Due to the relative comparability of CPU performance between the kernel and eBPF implementations of TCP Cubicm, we theorize the gap between eBPFCCP and kernel module CCP was due to our choice of using unix pipes rather than netlink as a channel for communicating between the eBPFCCP and the user space program. \\
From these results, we can conclude that eBPF is a viable platform for the development CCAs. The negligible performance overhead when compared to the kernel native implementation is impressive, and suggests that eBPF-CCAs can possibly also be used as a long-term deployment option outside of research and development stages. With regards to CCP, while the performance overhead of eBPFCCP is non-negligible, it should be noted that the benchmark is performed on a single core VM. In a real-world scenario, the performance overhead of eBPFCCP is likely to be negligible for all but the most performance-sensitive industry networking applications. The safety guarantees provided by eBPFCCP would provide a good intermediate step for users to adopt and experiment with a new CCA before it is natively implemented either directly into the kernel or as an eBPF program. 
\end{document}

>>>>

\subsection{Setup and Method}
An alternative implementation of an existing CCA can be evaluated using two metrics.
We evaluate eBPFCCP using two metrics:
\begin{enumerate}
    \item {\bf Behavior}: how accurate is eBPFCCP in replicating original CCA behavior?
    \item {\bf Performance}: what is the performance overhead of eBPFCCP?
\end{enumerate}

We included two sets of benchmarks to measure this: (1) a behavioral benchmark aimed to verify the "correctness" of different implementations (TCP Cubic should behave similarly no matter where it is implemented, as indicated by metrics such as bitrate/congestion window size over time), and (2) a performance benchmark aimed to measure the CPU overhead of different implementation approaches. A CCA implementation with smaller CPU overhead would logically lead to a higher maximum bandwidth in an ideal, infinite-speed connection where the CPU running the CCA is the bottleneck. \\
Our benchmarks utilize iperf3 \cite{iperf}. For each trial, we loop through every CCA we are testing and run iperf3 for $15$ seconds. At the same time, we run a Python script that loops infinitely while incrementing a variable. Once the $15$ seconds are up, the script then outputs the final number into a file. A higher output number suggests that the script went through more CPU cycles during the trial, indicating better CPU performance. We a ran a total of $20$ trials. \\
After all the trials, we average the behavioral metric across the trials and plot the average for each CCA through time as the behavioral benchmark. We then plot the CPU performances for each CCA across trials on a boxplot. \\
To make sure we are correctly benchmarking the performance impact of CCAs rather than miscellneous confounding factors, we ran the benchmark on an out-of-the-box Debian image on a VM. To ensure the CPU benchmark and the CCA are running on the same core, our VM is set up to use only one core. We also decided to run our benchmarks on various implementations of TCP Cubic due to its wide availability across platforms.

@@

The results of the behavioral benchmarks are shown in Figures~\ref{fig:cca-behavior-cwnd},~\ref{fig:cca-behavior-bitrate}, and~\ref{fig:cca-behavior-rtt}. From the graphs, we can see that the behavior of the same Congestion Control Algorithm remains similar across congestion window size, bitrate, and RTT. The results of the CPU benchmarks are shown in Figure \ref{fig:cca-performance}. The eBPF implementation of TCP Cubic achieved similar CPU performance compared to the built-in Linux kernel networking stack implementation. CCP-Cubic from the CCP Project \cite{generic-cong-avoid} combined with the CCP datapath kernel module and netlink slightly lower performance, where its median is at x percent of the full kernel implementation value. Our implementation of eBPFCCP ranked lower on the performance chart, but was still able to achieve y percent of the full kernel implementation value. Due to the relative comparability of CPU performance between the kernel and eBPF implementations of TCP Cubicm, we theorize the gap between eBPFCCP and kernel module CCP was due to our choice of using unix pipes rather than netlink as a channel for communicating between the eBPFCCP and the user space program. \\
From these results, we can conclude that eBPF is a viable platform for the development CCAs. The minor performance overhead when compared to the kernel native implementation suggests that eBPF-based CCAs can be used as a long-term deployment option outside of research and development stages. With regards to CCP, while the performance overhead of eBPFCCP is non-negligible, it should be noted that the benchmark is performed on a single core VM. In a real-world scenario, the performance overhead of eBPFCCP is likely to be negligible for all but the most performance-sensitive industry networking applications. The safety guarantees provided by eBPFCCP would provide a good intermediate step for users to adopt and experiment with a new CCA before it is natively implemented either directly into the kernel or as an eBPF program. 
\end{document}

### 2024-12-12 paper/sections/introduction.tex
paper: revise

\documentclass[../main.tex]{subfiles}
\begin{document}
TCP congestion control forms a critical part of the Internet ecosystem. It is often an active area of research, with new algorithms being proposed to improve network performance \cite{CCA-BBR, CCA-FCC} and discussions about how CCAs influence fairness between users \cite{CCAStarvation, CCA-Fairness}. Hence, congestion control research is critical to the stable operation and continuous improvement of the Internet, even in a post-flow rate fairness era ~\cite{PrinciplesForCongestionManagement, Dismantling}.\\
A key part of congestion control research involves developing and experimenting with new CCAs. Traditionally, this poses two main problems:
\begin{enumerate}
    \item {\bf Development friction}: CCA implements are often split between different datapaths. In the Linux kernel, this involves writing kernel code or modules, which presents a high barrier of entry and is prone to bugs that can crash the system.
    \item {\bf Distribution difficulties}: Pushing a new CCA upstream to the Linux kernel is difficult. To see real-world adoption, a new CCA must be user-friendly and easy to adopt in the early stages. Being easy to adopt is often predicated on ensuring safety and performance guarantees.
\end{enumerate}
To accurately evaluate the impacts of a new CCA on the Internet, it requires significant real-world adoption from users around the world. As a result, any CCA wishing to see eventual large-scale adoption must first propose a relatively user-friendly way for people to adopt it in the early stages. \\
With recent advancements in the eBPF space, the possibility of leveraging eBPF as a platform for implementing CCAs has emerged. eBPF \cite{ebpf} can be used to write and install programs that run in the Linux kernel with certain safety guarantees. Using the \texttt{struct\_ops} API, congestion control-specific hooks can be attached to the TCP stack, allowing for the implementation of new CCAs. This approach has the potential to solve both the development friction and distribution difficulties of developing new CCAs. The eBPF verifier provides powerful guarantees to the safety of the program, protecting against large amounts of hard-to-solve bugs that lead to the kernel crashing. Furthermore, the runtime-attached nature of eBPF makes iterating and testing new versions much easier. On the adoption side, the safety guarantees provided by eBPF will greatly encourage user adoption. While users can be hesitant in installing kernel patches or modules, attaching a eBPF program can be much lower-risk and appealing. \\
Despite these upsides, running CCAs in eBPF can possibly incur extra costs due to running in the eBPF VM. By instrumenting the performance overheads of various different ways of implementing CCAs, this paper explores the degree to which eBPF can be used as a platform for developing CCAs. In addition to CCAs directly implemented using eBPF, other means of developing CCAs quickly and safely are also tested.The Congestion Control Plane (CCP) \cite{cpp}, for example, is an idea proposed to implement CCAs in user space. Since the original CCP kernel datapath is written as a Linux kernel module, we also implemented an alternative datapath in eBPF to potentially take advantage of the safety guarantees.
\end{document}

>>>>

\documentclass[../main.tex]{subfiles}
\begin{document}
TCP congestion control forms a critical part of the Internet ecosystem. It is an active area of research that involves new algorithms being proposed to improve network performance \cite{CCA-BBR, CCA-FCC} and discussions about how CCAs influence fairness between users \cite{CCAStarvation, CCA-Fairness}. Hence, congestion control research is critical to the stable operation and continuous improvement of the Internet, even in a post-flow rate fairness era ~\cite{PrinciplesForCongestionManagement, Dismantling}.\\
A key part of congestion control research involves developing and experimenting with new CCAs. Traditionally, this poses two main problems:
\begin{enumerate}
    \item {\bf Development friction}: CCA implements are often split between different datapaths. In the Linux kernel, this involves writing kernel code or modules, which presents a high barrier of entry and is prone to bugs that can crash the system.
    \item {\bf Distribution difficulties}: Pushing a new CCA upstream to the Linux kernel is difficult. To see real-world adoption, a new CCA must be user-friendly and easy to adopt in the early stages. Being easy to adopt is often predicated on ensuring safety and performance guarantees.
\end{enumerate}
To accurately evaluate the impacts of a new CCA on the Internet, it requires significant real-world adoption from users around the world. As a result, any CCA wishing to see eventual large-scale adoption must first propose a relatively user-friendly way for people to adopt it in the early stages. \\
With recent advancements in the eBPF space, the possibility of leveraging eBPF as a platform for implementing CCAs has become more apparent. eBPF \cite{ebpf} can be used to write and install programs that run in the Linux kernel with certain safety guarantees. Using the \texttt{struct\_ops} API, congestion control-specific hooks can be attached to the TCP stack, allowing for the implementation of new CCAs. This approach has the potential to solve both the development friction and distribution difficulties of developing new CCAs. The eBPF verifier provides powerful guarantees to the safety of the program, protecting against large amounts of hard-to-solve bugs that may lead to kernel crashes. Furthermore, the runtime-attached nature of eBPF makes iterating and testing new versions much easier. On the adoption side, the safety guarantees provided by eBPF would greatly encourage user adoption. While users can be hesitant in installing kernel patches or modules, attaching a eBPF program can be much lower-risk and thus appealing. \\
However, running CCAs in eBPF can possibly incur extra costs due to running in the eBPF VM. By considering the performance overheads of various different ways of implementing CCAs, this paper explores the degree to which eBPF can be used as a platform for developing CCAs. In addition to CCAs directly implemented using eBPF, other means of developing CCAs quickly and safely are also tested.
\end{document}

### 2024-12-12 paper/sections/relatedwork.tex
paper: revise

\documentclass[../main.tex]{subfiles}
\begin{document}
CCP decoupled data gathering and CCA computation. A datapath module gathers information regarding a TCP flow and sends it to user space via various methods (such as netlink sockets and unix pipes), while a user space program listens to the given data and sends back any updates to be made to the TCP connection state. Platform/OS independence
\end{document}

>>>>

\documentclass[../main.tex]{subfiles}
\begin{document}
In order to fully evaluate performance of eBPF-based CCAs against alternative approaches, we looked at prior research aimed at reducing CCA-development overhead. Specifically, CCP \cite{ccp} was developed to be an abstraction layer where developers can experiment with new CCA ideas without worrying about the intricacies of specific datapath implementations. Most notably, CCP decoupled data gathering and CCA computation. A custom datapath module gathers TCP flow state and sends it to user space via various channels, such as netlink sockets and unix pipes. A user space program then listens to the provided data and sends back any updates to be made to the TCP connection state. Since the computation lives in user space, developers are shielded from the potential of introducing kernel-crashing bugs as long as the underlying CCP datapath implementation is safe. \\ 
Due to CCP's datapath-reliant nature, the performance of any CCP-based congestion control algorithm will depend on the implementation of the underlying datapath, as well as the channel used to communicate between user space and kernel space. In our investigations to test the performance of CCP-based CCAs, we built an alternative datapath module that is implemented as an eBPF program rather than a kernel module. In addition to providing more safety guarantees, this implementation also allows us to test how the performance of CCP-based CCAs can vary depending on the underlying datapath implementation. \\

\end{document}
